Julianna's algal metrics wrangling code

### 2024-08-15 julianna_algalMetric_wrangling.Rmd
Cleaned algal metric wrangling to have the same names as the datasets shared for folks to use with their ancillary datasets

@@

```{r join}
algDiv %>% 
  full_join(betaDHabYr) %>% 
  # follow a cleaner naming convention
  rename(algal_perc_cov = AlgCov_perc,
         algal_perc_cov_hard = AlgCov_percHard,
         algal_richness = Richness,
         algal_Shann = Shannon) %>% 
  # get just the columns we need
  select(-site_habitat, -hab_year, -island_side_hab_year) -> quadData
```

@@

```{r get averages}
quadData %>% 
  group_by(year, site, habitat, island_side) %>% 
  summarize(mean_coral_perc_cov = mean(coral),
            mean_alg_perc_cov = mean(algal_perc_cov),
            mean_alg_perc_cov_hard = mean(algal_perc_cov_hard),
            mean_alg_richness = mean(algal_richness),
            mean_alg_Shann = mean(algal_Shann),
            mean_dist_centroid_habyear = mean(hab_yr_distcentroid, na.rm = TRUE)
            ) -> siteData
```

@@

```{r calculate CV and summarize}
quadData %>% 
  # don't include year this time
  group_by(site, habitat, island_side) %>%
  summarize(
    mean_coral_perc = mean(coral),
    mean_alg_richness = mean(algal_richness), 
    richness_se = sd(algal_richness)/sqrt(n()),
    richness_cv = CV(algal_richness),
    mean_Shann = mean(algal_Shann), 
    Shann_se = sd(algal_Shann)/sqrt(n()),
    Shann_cv = CV(algal_Shann),
    # for total algal cover
    mean_alg_cov = mean(algal_perc_cov),
    alg_cov_se = sd(algal_perc_cov)/sqrt(n()),
    cv_alg_cov = CV(algal_perc_cov),
    # and for % of hard substrate
    mean_alg_cov_hard = mean(algal_perc_cov_hard),
    alg_cov_hard_se = sd(algal_perc_cov_hard)/sqrt(n()),
    cv_alg_cov_hard = CV(algal_perc_cov_hard),
    ) -> alltimeCV
```

```{r save}
# write_csv(alltimeCV, here(here("data/algal_metrics_all_time.csv")))
```

- Noam: Add **synchrony** to this dataset + any other relevant metrics

s

>>>>

This is for if you're not running betadisper() for the first time (loading old runs)

@@

```{r join}
algDiv %>% 
  full_join(betaDHabYr) %>% 
  # follow a cleaner naming convention to match Noam's
  rename(algae_cover = AlgCov_perc,
         algae_cover_hard = AlgCov_percHard,
         algae_richness = Richness,
         algae_shannon = Shannon) %>% 
  # get just the columns we need
  select(-site_habitat, -hab_year, -island_side_hab_year) -> quadData

```

@@

```{r get averages}
quadData %>% 
  group_by(year, site, habitat, island_side) %>% 
  summarize(coral_cover_mean = mean(coral),
            N = n(),
            coral_cover_se = sd(coral)/sqrt(N),
            algae_cover_mean = mean(algae_cover),
            algae_cover_se = sd(algae_cover)/sqrt(N),
            algae_cover_hard_mean = mean(algae_cover_hard),
            algae_cover_hard_se = sd(algae_cover_hard)/sqrt(N),
            algae_richness_mean = mean(algae_richness),
            algae_richness_se = sd(algae_richness)/sqrt(N),
            algae_shannon_mean = mean(algae_shannon),
            algae_shannon_se = sd(algae_shannon)/sqrt(N),
            centroid_dist_habyear_mean = mean(hab_yr_distcentroid, na.rm = TRUE)
            ) -> siteData

```

@@

```{r calculate CV and summarize}
quadData %>% 
  # don't include year this time
  group_by(site, habitat, island_side) %>%
  summarize(
    coral_cover_mean = mean(coral),
    algae_richness_mean = mean(algae_richness), 
    algae_richness_se = sd(algae_richness)/sqrt(n()),
    algae_richness_cv = CV(algae_richness),
    algae_shannon_mean = mean(algae_shannon), 
    algae_shannon_se = sd(algae_shannon)/sqrt(n()),
    algae_shannon_cv = CV(algae_shannon),
    # for total algal cover
    algae_cover_mean = mean(algae_cover),
    algae_cover_se = sd(algae_cover)/sqrt(n()),
    algae_cover_cv = CV(algae_cover),
    # and for % of hard substrate
    algae_cover_hard_mean = mean(algae_cover_hard),
    algae_cover_hard_se = sd(algae_cover_hard)/sqrt(n()),
    algae_cover_hard_cv = CV(algae_cover_hard),
    ) -> alltimeCV
```

Combine with Noam's synchrony/portfolio estimates:

```{r}
noamallTime <- read_csv(here("data/algae_diversity_stability_metrics_alltime.csv"))
```

```{r}
noamallTime %>% 
  select(site, habitat, island_side, algae_shannon_stability, algae_loreau_mean, algae_loreau_se, algae_portfolio_mean, algae_portfolio_se) %>% 
  full_join(alltimeCV) -> alltimeJoin
```

```{r save}
write_csv(alltimeJoin, here(here("data/algal_metrics_all_time.csv")))
```
